---
title: "Explore gapseq biochemistry"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Explore gapseq biochemistry}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Before you start

This package (gapseq.biochem) is a small tool to explore the biochemistry
database behind gapseq. It is *not* the gapseq software itself. You can find all
information on gapseq here:

- Publication: [Genome Biology](https://doi.org/10.1186/s13059-021-02295-1)
- Documentation: [ReadTheDocs](https://gapseq.readthedocs.io/en/latest/)
- Source code: [GitHub](https://github.com/jotech/gapseq)

***Important note:*** The biochemistry database in gapseq is derived from 
ModelSEED (https://modelseed.org/). In order to give credits to the developers
and curators of ModelSEED, we strongly encourage you to cite also the
publication for the [ModelSEED database](https://doi.org/10.1093/nar/gkaa746)
when you are using gapseq.

## Initialization

With the function `initDB()`, the database can be initialized from different 
sources and or different versions.

```{r setup}
library(gapseq.biochem)

# Latest version from github
db <- initDB()

# Version at a specific commit
#db <- initDB("c1761254bf65f2789e012fd6ae2da3e45f2b6bae") 

# Database from a local gapseq installation
#db <- initDB("~/Software/gapseq/")
```

## Exploring the biochemistry database

### Summary of a specific reaction
```{r}
summary(db, "rxn10132")
```

### Summary of a specific compound

Example: Summary for Lactose
```{r}
summary(db, "cpd00208")
```

### Find reactions
There are different ways to search for specific reactions, e.g. via specific
compound IDs (if they are known), via regular expressions/search terms for
compound names and/or by EC-numbers. Everything can be done with a single
function: `findRxns()`. Here are a few example cases:

Example: Find reactions with the compound cpd00211 (Butyrate)
```{r}
findRxns(db, cpd = "cpd00211")
```

Example: Find reactions with the both compound ids: cpd00100 (Glycerol) and 
cpd22614 (sulfoquinovose)
```{r}
findRxns(db, cpd = c("cpd00100","cpd22614"))
```

Example: Find reactions with at least one of the compound ids: cpd22614
(sulfoquinovose) or cpd01337 (Romicil).
```{r}
findRxns(db, cpd = c("cpd22614","cpd01337"), cpd.link = "OR")
```

Example: Find reactions for EC 4.2.1.22.
```{r}
findRxns(db, ec = "4.2.1.22")
```

Example: Find reactions, where the on participating metabolite contains the term
"sulfoquinovose"
```{r}
findRxns(db, grep.str = "sulfoquinovose")
```

Example: Find all reactions that include NADH (cpd00008) as co-factor and where
at least one compound matches the term "L-Lactate" or "D-Lactate":
```{r}
findRxns(db, cpd = "cpd00004", grep.str = "(D|L)-Lactate")
```



